Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data

### 2022-03-22 Manuscript/R&R/00-response_letter.Rmd

---
output: 
  pdf_document:
    keep_tex: FALSE
bibliography: reply_ref.bib
header-includes: 
 - \usepackage{color}
link-citations: true
---

\newcommand{\bg}[1]{\textcolor{red}{#1}}

@@

\bigskip

# Response to Associate Editor
_Comments to the authors:_

_Please follow carefully the comments, criticisms and suggestions of the three referees._

__Response__: We summarize the major changes of our manuscript as a response to address the comments from the referees:
\begin{itemize}
    \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects to reflect the effect hierarchy principle for bi-level selection. We updated the simulation and real data analysis results.
    \item We streamlined the structure of the manuscript, moving the previous EM-IWLS fitting algorithm subsection to the supporting information for three reasons. First of all, moving can help the audience focus on the model formulation and EM-Coordinate Descent algorithm, avoiding confusion and distraction. Secondly, many useful comments raised by the reviewers related to the EM-IWLS algorithm, e.g. constructing hypothesis testings and confidence band, requires a more dedicated discussion. We are preparing another manuscript discussing those. Lastly, with the added variable selection performance, the manuscript would be length if keeping the EM-IWLS in the main body. The readability of the manuscript would be damaged. Relocating the EM-IWLS section solves the dilemma without hurting the integrity of the current manuscript that emphasize the scalability.
    \item According to Reviewer \#1's and Reviewer \#4's comments, we added linear LASSO model and spikeSlabGAM model among the models to be compared in the simulation study. In addition, we added a new set of simulation scenarios where the functions of predictors are linear.  
    \item According to Reviewer \#4's comment, we summarized the variable selection and bi-level selection performance in the simulation study section. The computation performance of spikeSlabGAM is recorded and addressed too.
    \item According to Reviewer \#4's comment, we added the plot showing the estimated functions of predictors for real-data analysis.
    \item We updated the feature screening procedure in our real-data analysis of Emory Cardiovascular Biobank dataset. Our previous screening method, fitting univariate GAM, made the "double-dipping" mistake that would overestimate the prediction performance of any downstream prediction model. We updated the screening procedure with variance-based selection, and the prediction of both methods dropped. Nevertheless, the proposed model still outperforms the compared method. See change on P9.
    \item We fixed the additional typos and grammar errors according to Review \#2. The revision of the manuscript is marked in \bg{red} (roughly due to the technical difficulties associated with Latex). Deletions are not highlighted.
\end{itemize}

### 2022-03-22 Manuscript/R&R/01-reviewer_1.Rmd
Revise response letter

__Response__: Thank you for the comment. We expanded our simulation study from two perspectives: 1) we added linear lasso models and spikeSlabGAM models, where linear lasso models serve as the reference to benchmark the performance of linear models; 2) we added new scenarios where the the functional form of each predictors are linear, to demonstrate the robustness of the proposed, regarding prediction and variable selection performances, when the non-linear assumption no longer holds. Please see the changes and conclusion in \emph{Section 3: SIMULATION STUDY}. To conclude, we observed [TODO: REPLACE HERE WITH CONCLUSION]. Hence, the proposed models perform better than linear models when functional are non-linear and as good as linear model when functions are linear. The expanded simulation study also demonstrates that the proposed model is flexible to model both linear and non-linear effects and broadly applicable in high-dimensional modeling without much constraints, which deems to be an additional strength.

@@

__Response__: We thank the reviewer for the comment. We didn't perform the data processing step ourselves. We used the publicly available data that has been pre-processed. The data processing step is described in [TODO: add citation]. Meanwhile, we respectively argue that the small R2 is reasonable, as in many applications, genomics data can only accounts small amount of data variation. We think this example would be the case, as environmental factors (not accounted in these models) would be more substantial when explaining weight loss.
\bigskip

@@

__Response__: Thank you for the comment. The purpose of mentioning the other several priors in the previous manuscript is to describe the generalization of the proposed algorithms. In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalably model. We included an additional sentence in the discussion section. [TODO: list the sentence here].

@@

We revised our method section to highlight these advantages, see Pxxx.

@@

__Response__: We don't think any preliminary checkings or testings are necessary before implementing the proposed model, as long as some general rules of thumb for high-dimensional modeling are followed. Such rules of thumb includes [TODO: Insert citaiton for ultra high-dimension thing, as the number of predictors is a function of the sample size.]. As a supporting evidence, our added simulation demonstrates the proposed model performs well, in reference to the linear lasso model, even when signals are linear and sparse (see response to the Comment 1).

We acknowledge the reviewer's comment that careful deliberation are necessary when conducting analysis, which includes domain knowledge. However, the discussion of modeling strategy, e.g. starting with more flexible models or linear models, or using statistical testings to examine assumptions, is out of the scope of the current manuscript. The current manuscript focuses on introducing a novel model that allows flexibility when modeling complex signals.

>>>>

__Response__: Thank you for the comment. We added linear LASSO models to our simulation studies and see that the proposed model predicts as good as linear LASSO models when the underlying signals are linear, and substantially better when the underlying signals are nonlinear. This empirically demonstrates the proposed model is flexible and does not need additional metrics to decide if the model is appropriate to model linear or non-linear signals. The simulation results also match with our expectation. The proposed model, by design, can automatically detect if the underlying is linear or nonlinear, framed as bi-level selection problem in the manuscript. When the underlying signal is linear, the nonlinear component of the smoothing function would not be included in the model due to the proposed two-part spike-and-slab lasso prior.

<!-- expanded our simulation study from two perspectives: 1) we added linear lasso models and spikeSlabGAM models, where linear lasso models serve as the reference to benchmark the performance of linear models; 2) we added new scenarios where the the functional form of each predictors are linear, to demonstrate the robustness of the proposed, regarding prediction and variable selection performances, when the non-linear assumption no longer holds. Please see the changes and conclusion in \emph{Section 3: SIMULATION STUDY}. To conclude, we observed [TODO: REPLACE HERE WITH CONCLUSION]. Hence, the proposed models perform better than linear models when functional are non-linear and as good as linear model when functions are linear. The expanded simulation study also demonstrates that the proposed model is flexible to model both linear and non-linear effects and broadly applicable in high-dimensional modeling without much constraints, which deems to be an additional strength. -->

@@

__Response__: Thank you for the comment. We didn't perform the data processing step ourselves. We used the publicly available data that has been pre-processed, which is clarified in the manuscript on P9.

<!-- %Meanwhile, we respectively argue that the small R2 is reasonable, as in many applications, genomics data can only accounts small amount of data variation. We think this example would be the case, as environmental factors (not accounted in these models) would be more substantial when explaining weight loss.  -->
\bigskip

@@

__Response__: Thank you for the comment. The purpose of mentioning the other several priors in the previous manuscript is to describe the generalization of the proposed algorithms. In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalable model. We included an additional sentence in the discussion section (see P10 "Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.").

@@

@@

__Response__: We thank the review for the comments. For prediction purpose, the proposed model can be flexibly used in the high-dimensional data analysis without any checking or testing. The two-part spike-and-slab LASSO prior automatically detects if the underlying signal is linear or complex. When the signal is linear, the smoothing function of the predictor only includes the linear part; when the signal is complex, cubic spline (implemented in the package) is used to estimate the complex signal. As a supporting evidence, our added simulation demonstrates the proposed model predicts as good as, if not better than, the linear LASSO model, even when signals are linear and sparse (see response to the Comment 1).

For the bi-level selection purpose, we agree with the reviewer that it is a more delicate problem. In our simulation study, we observe that trade-off exists with different underlying signals compared to the only other method, spikeSlabGAM, that is capable of bi-level selection. (see P8) We added a discussion about the reason for the trade-off and gave the reader suggestions (see P10).

In conclusion, the proposed model do not addition detection for prediction purpose, and have different utility for bi-level selection.

<!-- reviewer's comment that careful deliberation are necessary when conducting analysis, which includes domain knowledge. However, the discussion of modeling strategy, e.g. starting with more flexible models or linear models, or using statistical testings to examine assumptions, is out of the scope of the current manuscript. The current manuscript focuses on introducing a novel model that allows flexibility when modeling complex signals. -->

<!-- We don't think any preliminary checking or testings are necessary before implementing the proposed model, as long as some general rules of thumb for high-dimensional modeling are followed. -->
<!-- Such rules of thumb includes [TODO: Insert citaiton for ultra high-dimension thing, as the number of predictors is a function of the sample size.].  -->

<!-- We acknowledge the reviewer's comment that careful deliberation are necessary when conducting analysis, which includes domain knowledge. However, the discussion of modeling strategy, e.g. starting with more flexible models or linear models, or using statistical testings to examine assumptions, is out of the scope of the current manuscript. The current manuscript focuses on introducing a novel model that allows flexibility when modeling complex signals. -->

### 2022-03-22 Manuscript/R&R/02-reviewer_2.Rmd
Revise response letter

__Response__:
\bigskip

@@

__Response__:
\bigskip

@@

__Response__:
\bigskip

@@

__Response__:
\bigskip

>>>>

__Response__: We thank the reviewer for the comment. The purpose of the reparameterization step is to factor the smoothing penalty matrix in the design matrix, such that the smoothing penalty can be addressed implicitly. This contrasts to some previous methods that do not consider smoothing penalty at all. With the reparameterization, the smoothing penalty changes from $\lambda \bs \beta^T \bs S \bs \beta$ to $\lambda \bs {\beta^*}^T \bs {\beta^*}$, where the $\bs \beta$ and $\bs \beta^*$ are the coefficients of the smoothing function before and after the reparameterization respectively. However, the smoothing parameter $\lambda$, equivalently the scale parameter $s$ in the proposed model, needs to be estimated/tuned. The proposed two-part spike-and-slab LASSO prior provides the the estimation of the smoothing parameter: when the nonlinear effect does not exists, a large shrinkage is applied to the coefficients via the spike density such that coefficients shrink to 0s; when then nonlinear exists, a small shrinkage is applied via the slab density such that the smoothing function can be estimated appropriately. Moreover, the local adaptivity of he spike-and-slab LASSO prior allows the shrinkage to adapt to the data, compared to previous methods that impose an uniform shrinkage.

We clarified the utility of the reparameterization step and emphasized how the smoothness of the additive function are appropriately modeled (see P3).

@@

__Response__: Compared to other priors, the spike-and-slab double exponential prior provides three advantages. First of all, the spike-and-slab double exponential prior provides a locally adaptive shrinkage when estimating the coefficients. Hence, the smoothing functions can be estimated more accurately. Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straight forward. Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.

We revised our method section to highlight these advantages, see P4.
\bigskip

@@

__Response__: We thank the reviewer for the comment. We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$, in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously. This prior formulation is motivated by effect hierarchy discussed in Chipman (2004). Now, the inclusion of the nonlinear effects is dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection. The modification of the prior doesn't complicate the model fitting algorithm. The $\gamma$ can be analytically integrated out of the density function of $\gamma^*$ and requires minimum change of the proposed algorithms. The updated model still shows performance advantages as before.

@@

__Response__: We thank the review for the comment. We revised the manuscript according to the comment.
\bigskip

### 2022-03-22 Manuscript/R&R/03-reviewer_4.Rmd
Revise response letter

__Response__:
\bigskip

@@

__Response__:
\bigskip

@@

__Response__:
\bigskip

@@

__Response__:
\bigskip

@@

__Response__:
\bigskip

@@

Nevertheless, we edited the sentence for clarity. [TODO: check if this is done]
\bigskip

>>>>

__Response__: We thank the reviewer for the comment. We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$, in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously. This prior formulation is motivated by effect hierarchy discussed in Chipman (2004). Now, the inclusion of the nonlinear effects is strictly dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection. The modification of the prior doesn't complicate the model fitting algorithm. The $\gamma$ can be analytically integrated out of the density function and requires minimum change of the proposed algorithm. In other words, the proposed model remain feasible for high-dimensional data analysis and retains the claimed computational advantages.
\bigskip

@@

__Response__: We thank the review for the comments. We made change to our simulation study section, mainly adding variable selection and a new set of simulations that the predictors are all linear. The suggested model spikeSlabGAM and linear Lasso model as the benchmark were added to the models of comparison. When reporting the variable selection performance, we considered three metrics, positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient. Bi-level selection performance was provided for the two applicable models, the proposed model and spikeSlabGAM. Please see the updated simulation study on P7-9.

@@

__Response__: We thank the review for the comment. In the update manuscript, we add the suggested model, spikeSlabGAM, in the simulation study, where all the performance metrics and computation time are recorded and compared. We see that the proposed method still preserves the computation advantage over spikeSlabGAM, particularly in the high-dimensional cases.
\bigskip

@@

__Response__: We thank the reviewer for the comment. After deliberation, we decide to focusing explaining the EM-CD method well. The inclusion of EM-IWLS would distract the reader from the main focus of manuscript, scalable model, good prediction and bi-level selection. We fully agree that the questions raised by the reviewer are very important and deserves more careful discussion. Nevertheless, that would be out of the scope of the current work. We will dedicate another manuscript to detail many of the questions here, for example estimating effective degree of freedom, limiting distribution of the test statistics to the EM-IWLS method.

@@

__Response__: We thank the reviewer for the comment. We added Figure 2 displaying the estimated functions of the predictors from BHAM.
\bigskip

@@
